Sketch of DP - Bayesian Inference

### 2018-07-16 main-sketch.tex
introduction to bayesian things

In this work we present a new differentially private algorithm to compute the posterior.
The noise is calibrated w.r.t. a smooth sensitivity computed with
a metric on probability distributions; we measured the accuracy
of the process by using a metric over probability measures.
We compared the accuracy of this approach with the one
based on $\ell_1$-norm. Experimental results show that
when the data size is small or when the parameters of the prior
distribution are large, the former outperforms the latter.

@@

\section{Problem Statement}
Our research is based on an observation that under different measurement, mechanisms usually have different accuracy behavior. When adding noise according to a certain measurement, mechanism' accuracy behavior will be better than when adding noise according to other measurements. Here we study the private Bayesian inference based on Hellinger distance measurement. Showing that our newly-designed mechanism adding noise according to Hellinger distance performs better than baseline mechanism adding noise according to $l_1$ norm.

\section{Technical Problem Statement}
\label{sec_bayesInfer}

We are studying the Bayesian inference in Beta-Bernoulli systems. So this part we will technically describe our problem under Beta-Bernoulli systems of Bayesian inference. Suppose $\pi(\xis)$ is our prior belief about some parameters or latent variables $\xis$. By using a group of observed data $\textbf{x}$ sizing $n$, the likelihood of this group of data is $Pr(\textbf{x} | \xis)$, we produce a posterior distribution:
\begin{equation*}
Pr(\xis | \textbf{x}) = \frac{Pr(\textbf{x} | \xis) \pi(\xis)}{Pr(\textbf{x})},
\end{equation*}
denoted as $\bysinfer(\textbf{x},\pi(\xis))$ taking an observed data set $\textbf{x} \in \mathcal{X}^n$ (i.e., $\{0,1\}^n$) and a prior distribution $\pi(\xis)$ as input, outputting a posterior distribution of the parameter $\alphas$. For conciseness, when prior is given, we use $\bysinfer(x)$. $n$ is the size of the observed data size.

In Beta-Bernoulli systems, we take a Beta distribution as prior belief for our parameters and latent variables, $\betad(\alpha, \beta)$, where $\pi(\xi) = \betad(\xi | \alpha, \beta) = \frac{\xi^{\alpha} (1 - \xi)^{\beta}}{B(\alphas)}$, and the Bernoulli distribution as the likelihood for $Pr(\textbf{x} | \xi) = \xi^{\sum\limits_{\mathrm{x_i} \in \textbf{x}}\mathrm{x_i}}(1 - \xi)^{n -\sum\limits_{\mathrm{x_i} \in \textbf{x}}\mathrm{x_i}} $. %We are throwing an irregular coin with side $A$ and side $B$. The posterior is computed as $(\alpha + n_1, \beta + n_0)$, where $n_1$ is the appearance time of side $A$ in the observed data set and $n_0$ is the appearance time of the other side.

This can be extended to Dirichlet distribution of $m$ dimensions $\dirichlet(\alphas)$ where $x$ take value from universe $\mathcal{X}$ of size $m$ (i.e., $\mathcal{X} = \{x_1, \cdots, x_m\}$). $\pi(\xis) = \dirichlet(\xis | \alphas) = \frac{\prod_{i = 1}^{m}\xi_i^{\alpha_i}}{B(\alphas)}$, and the likelihood for $Pr(x | \xis)$ is $\prod\limits_{\xi_i \in \xis}\xi_i^{\sum\limits_{\{\mathrm{x}_i|\mathrm{x}_i = x_i\}} \mathrm{x_i}}$.

% We give a inference process based on a concrete example, where we throw an irregular $m$ sides dice. We want to infer the probability of getting each side $\xis$. We get a observed data set $\{s_{k_1}, s_{k_2}, \cdots, s_{k_n}\}$ by throwing the dice $n$ times, where $k_i \in \{1,2,\cdots, m\}$ denotes the side we get when we throw the dice the $i^{th}$ time. The posterior distribution is still a Dirichlet distribution with parameters $(\alpha_1 + n_1, \alpha_2 + n_2, \cdots, \alpha_m + n_m)$, where $n_i$ is the appearance time of the side $i$ in total.

% In the case that $m = 2$, it is reduced to a Beta distribution $\betad(\alpha, \beta)$. The $m$ side dice change into a irregular coin with side $A$ and side $B$. The posterior is computed as $(\alpha + n_1, \beta + n_0)$, where $n_1$ is the appearance time of side $A$ in the observed data set and $n_0$ is the appearance time of the other side.

The posterior belief about parameters $\xis$ is inferred from the group of sensitive data. Because it is usually published in form of a distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself can relieve a lot of information about sensitive data. So, the sensitive information we want to protect here is the posterior belief, i.e., $Pr(\xis | \textbf{x})$, rather than just a sample drew from it\cite{foulds2016theory}. Then it will be better if we can use a distance over two distributions to measure the accuracy than a distance over just parameters or values when we are protecting a distribution, for instance, Hellinger distance in our paper. When we measure the accuracy between true posterior and protected posterior using Hellinger distance, it would be better we can have a protection mechanism based on Hellinger distance. However, current mechanisms are mainly based on $l_1$ norm by Laplace mechanism \cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we design our exponential mechanism over Hellinger distance to protect the whole posterior distribution.

>>>>

In this work we present a new differentially private algorithm to
compute the posterior in a Beta-Binomial system and in Dirichlet-Multinomial system.
The noise is calibrated w.r.t. a smooth
sensitivity computed with a metric on probability distributions; we
measured the accuracy of the process by using a metric over
probability measures. We compared the accuracy of this approach with
the one based on $\ell_1$-norm. Experimental results show that when
the data size is small or when the parameters of the prior
distribution are large, the former outperforms the latter.

@@

\section{Bayesian Inference Background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{Pr(\dataobs | \theta) \cdot \Pr(\xis)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
\dataobs~ is observed. Since we consider \dataobs to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias})--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{a}(1-\theta)^{n-a}
\]
where $a=\displaystyle\sum_{i=1}^{n}\mathrm{x_i}$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha+a,\beta+n-a)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\vtheta)$, for $k\in\mathbf{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a\in\datauni}\alpha_{a}^{z_a}
\todo{to check}\]
where $z_a=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\todo{to type down}\dirichlet(\vtheta+...)
\]
\section{Problem Statement}
% When we try
% to make this distribution public, sensitive information of the data
% based on which the posterior drew from can be compromised. So, it is
% important to protect this posterior distribution. When our protection
% objective is a distribution rather than a numeric value or vector, it
% would be better to measure the accuracy by distance over
% distributions, for instance, Hellinger distance, than by numeric
% metrics like $\ell_1$-norm.
This work will show that the same mechanism can
perform differently when its accuracy is computed w.r.t to a metric
over probability measures (e.g total variation distance, or Hellinger distance)
from when the metric (e.g. $\ell_1$-norm) is over the parameters.  
\section{Technical Problem Statement}
\label{sec_bayesInfer}
The posterior belief about parameters $\xis$ is inferred from the
group of sensitive data. Because it is usually published in form of a
distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself
can relieve a lot of information about sensitive data. So, the
sensitive information we want to protect here is the posterior belief,
i.e., $Pr(\xis | \dataobs)$, rather than just a sample drew from
it\cite{foulds2016theory}. Then it will be better if we can use a
distance over two distributions to measure the accuracy than a
distance over just parameters or values when we are protecting a
distribution, for instance, Hellinger distance in our paper. When we
measure the accuracy between true posterior and protected posterior
using Hellinger distance, it would be better we can have a protection
mechanism based on Hellinger distance. However, current mechanisms are
mainly based on $l_1$ norm by Laplace mechanism
\cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we
design our exponential mechanism over Hellinger distance to protect
the whole posterior distribution.
